---
title: |
  | Bayesian Modeling of Hurricane Trajectories
author: |
  | Hongjie Liu, Xicheng Xie, Jiajun Tao, Zijian Xu, Shaohan Chen
date: "May 1st, 2023"
header-includes:
   - \usepackage{bm}
   - \usepackage{graphicx}
   - \usepackage{float}
   - \usepackage{subfigure}
   - \usepackage{algorithm}
   - \usepackage{algpseudocode}
output:
  beamer_presentation:
    colortheme: "default"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(sigmoid) 
library(qgam) 
library(pROC)
library(xtable)
library(kableExtra)
library(boot) 
library(ggplot2)
library(gridExtra)
# magic that automatically adjusts the font size
def.chunk.hook = knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x = def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```


## Outline

- Introduction

- EDA

- Task 1

- Task 2

- Task 3

- Task 4

- Discussions

- Q&A


## Introduction
**Background:**

- A hurricane is a large and powerful tropical cyclone that typically forms over warm ocean waters and can cause significant damage and destruction to coastal areas.

**Motivation:**

- Researchers are interested in modeling hurricane trajectories to forecast wind speed to predict the severity or to develop protective measures.

## Introduction 
**Data Source:**

- "hurricane703.csv" collected the track data (every 6 hours) of 702 hurricanes in the North Atlantic area since 1950.

**Variables:**

- ID: ID of the hurricanes
- Season: In which the hurricane occurred
- Month: In which the hurricane occurred
- Nature: Nature of the hurricane
  
  ET: Extra Tropical
  
  DS: Disturbance
  
  NR: Not Rated

  SS: Sub Tropical
  
  TS: Tropical Storm

- time: dates and time of the record
- Latitude and Longitude: The location of a hurricane check point
- Wind.kt: Maximum wind speed (in Knot) at each check point

## EDA
**Data pre-processing:**

- We only kept observations that occurred on 6-hour intervals.

- We found that some hurricanes had the same ID but were actually different ones.

- We excluded hurricanes that had fewer than 3 observations.

- We defined August, September, and October as active season, the rest as inactive season.

- After the process, there are 21691 observations across 704 unique hurricanes.

## EDA
\begin{figure}[ht]
  \centering
  \includegraphics[width=1.0\textwidth]{jacky_part_pic/traj.png}
  \caption{Trajectories of Hurricanes by Month}
\end{figure}

## EDA 
\begin{figure}[ht]
  \centering
  \includegraphics[width=1.0\textwidth]{jacky_part_pic/month.png}
  \caption{Count and Average Speed of Hurricanes in each Month}
\end{figure}

## EDA 
\begin{figure}[ht]
  \centering
  \includegraphics[width=1.0\textwidth]{jacky_part_pic/year.png}
  \caption{Count and Average Speed of Hurricanes in each Year}
\end{figure}

## Bayesian Model
Let $Y_{i}(t)$ denote the wind speed of the $i$th hurricane at time $t$ (in hours) since the hurricane began. The following Bayesian model was suggested to model the wind speed of the $i$th hurricane 6 hours later:
$$\begin{aligned}
Y_{i}(t+6) =&\ \beta_{0,i}+\beta_{1,i}Y_{i}(t) + \beta_{2,i}\Delta_{i,1}(t)+
\beta_{3,i}\Delta_{i,2}(t) +\beta_{4,i}\Delta_{i,3}(t)  \\
&\ + \mathbf{X}_i^\top\boldsymbol\gamma+ \epsilon_{i}(t),\end{aligned}$$
where 

* $\Delta_{i,1}(t)$, $\Delta_{i,2}(t)$ and $\Delta_{i,3}(t)$: changes of latitude, longitude and wind speed between $t-6$ and $t$
  - random coefficients $\boldsymbol{\beta}_{i} = (\beta_{0,i},\beta_{1,i},...,\beta_{4,i})^\top$
* $\mathbf{X}_i = (x_{i,1},\ldots,x_{i,6})^\top$: covariates with fixed effects $\boldsymbol\gamma$, where
  - $x_{i,1}$: the calendar year of the $i$-th hurricane
  - $x_{i,2}$: indicator variable of the month in active season (August-October) when the $i$-th hurricane started
  - $x_{i,3},\ldots,x_{i,6}$: indicator variables of the type (ES, NR, SS, TS) of the $i$-th hurricane
* $\epsilon_{i,t}\sim N(0,\sigma^2)$, independent across $t$

## Task 1 - Prior Distributions

**Objective:** Let $\mathbf{B}= (\boldsymbol{\beta}_{1}^\top,...,\boldsymbol{\beta}_{n}^\top)^\top$, derive the posterior distribution of the parameters $\Theta=(\mathbf{B}^\top,\boldsymbol{\mu}^\top,\boldsymbol\Sigma,\boldsymbol\gamma^\top,\sigma)$.

We assume that

* $\boldsymbol{\beta}_{i} \overset{i.i.d.}{\sim} N(\boldsymbol{\mu}, \boldsymbol{\Sigma})$

* $\boldsymbol{\mu}\sim N(\boldsymbol{0},\boldsymbol{V})$

* $\boldsymbol{\Sigma}$: an inverse-Wishart distribution with d.f. $\nu$ and scale matrix $\boldsymbol{S}$

* $\boldsymbol\gamma\sim N(\boldsymbol 0,0.05^2\boldsymbol I_{6})$

* $\sigma$: a half-Cauchy distribution with scale parameter 10

We set $\boldsymbol{V} =\boldsymbol{S} = \boldsymbol I_{5}$, and $\nu=5$.

## Task 1 - Joint Prior Distribution of Parameters
Let $n$ denote the number of hurricanes in the dataset. The prior distribution of $\Theta=(\mathbf{B}^\top,\boldsymbol{\mu}^\top,\boldsymbol\Sigma,\boldsymbol\gamma^\top,\sigma)$ is given by
$$\begin{aligned}
\pi(\Theta)=&\ \pi(\mathbf{B}^\top,\boldsymbol{\mu}^\top,\boldsymbol{\Sigma},\boldsymbol\gamma^\top,\sigma)\\
=&\ \pi(\mathbf{B}^\top\mid\boldsymbol{\mu}^\top,\boldsymbol{\Sigma})\pi(\boldsymbol{\mu}^\top,\boldsymbol{\Sigma})\pi(\boldsymbol\gamma)\pi(\sigma)\\
=&\ \left(\prod_{i=1}^n\pi(\boldsymbol\beta_i^\top\mid\boldsymbol{\mu}^\top,\boldsymbol{\Sigma})\right)\pi(\boldsymbol{\mu})\pi(\boldsymbol{\Sigma})\pi(\boldsymbol\gamma)\pi(\sigma)\\
\propto &\ |\boldsymbol{\Sigma}|^{-n/2}\exp\left(-\frac{1}{2}\sum_{i=1}^n(\boldsymbol\beta_i-\boldsymbol\mu)^\top\boldsymbol{\Sigma}^{-1}(\boldsymbol\beta_i-\boldsymbol\mu)\right)\\
&\ \times\exp\left(-\frac{1}{2}\boldsymbol\mu^\top\boldsymbol{V}^{-1}\boldsymbol\mu\right)\\
&\ \times|\boldsymbol{\Sigma}|^{-(\nu+6)/2}\exp\left(-\frac{1}{2}\mathrm{tr}(\boldsymbol{S}\boldsymbol{\Sigma}^{-1})\right)\\
&\ \times \exp\left(-\frac{1}{2}\cdot400\boldsymbol\gamma^\top\boldsymbol\gamma\right)\times\frac{I(\sigma>0)}{1+(\sigma/10)^2}.
\end{aligned}$$

## Task 1 - Likelihood

Let $m_i$ denote the number of observations and $\boldsymbol{Y}_i=(Y_{i,1},\ldots Y_{i,m_i})^\top$ denote the wind speed data of the $i$-th hurricane (excluding the first and second observations), where $Y_{i,k}=Y_i(6k+6)$. Denote $\boldsymbol{Y}=(\boldsymbol{Y}_1^\top,\boldsymbol{Y}_2^\top,\ldots,\boldsymbol{Y}_n^\top)^\top$, and $\mathbf Z_{i,k}=(1,Y_{i,k},\Delta_{i,1}(6k+6),\Delta_{i,2}(6k+6),\Delta_{i,3}(6k+6))^\top$.

Given that
$$Y_{i,j}\mid (\boldsymbol\beta_i^\top,\boldsymbol\gamma^\top,\sigma)\sim N(\mathbf Z_{i,j-1}^\top\boldsymbol\beta_i + \mathbf{X}_i^\top\boldsymbol\gamma,\sigma^2),$$
we have
$$\begin{aligned}
L(\Theta\mid\boldsymbol{Y}^\top)=&\ \prod_{i=1}^n\prod_{j=1}^{m_i}L(\Theta\mid Y_{i,j})\\
=&\ \prod_{i=1}^n\prod_{j=1}^{m_i}\left[\frac{1}{\sqrt{2\pi}\sigma}\exp\left(-\frac{(Y_{i,j}-\mathbf Z_{i,j-1}^\top\boldsymbol\beta_i-\mathbf{X}_i^\top\boldsymbol\gamma)^2}{2\sigma^2}\right)\right].
\end{aligned}$$

## Task 1 - Joint Posterior Distribution of Parameters

$$\begin{aligned}
\pi(\Theta\mid \boldsymbol{Y}^\top)
\propto &\ L(\Theta\mid\boldsymbol{Y}^\top)\pi(\Theta)\\
\propto &\ \frac{I(\sigma>0)\sigma^{-\sum_{i=1}^n m_i}}{1+(\sigma/10)^2}|\boldsymbol{\Sigma}|^{-(n+\nu+6)/2}\\
&\ \times\exp\Bigg[-\frac{1}{2}\Bigg(\sum_{i=1}^n(\boldsymbol\beta_i-\boldsymbol\mu)^\top\boldsymbol{\Sigma}^{-1}(\boldsymbol\beta_i-\boldsymbol\mu)+\boldsymbol\mu^\top\boldsymbol{V}^{-1}\boldsymbol\mu\\
&\quad\quad\quad\quad\quad\quad\ +\mathrm{tr}(\boldsymbol{S}\boldsymbol{\Sigma}^{-1})+400\|\boldsymbol\gamma\|_2^2\Bigg)\Bigg]\\
&\ \times \exp\left(-\frac{1}{2\sigma^2}\sum_{i=1}^n\sum_{j=1}^{m_i}(Y_{i,j}-\mathbf Z_{i,j-1}^\top\boldsymbol\beta_i-\mathbf{X}_i^\top\boldsymbol\gamma)^2\right).
\end{aligned}$$


## Task 2
**Objective:**\par

Design and implement a custom MCMC algorithm for the outlined Bayesian hierarchical model. Monitor the convergence of the MCMC chains, using diagnostic plots and summary statistics to check for any issues.

## Task 2 - Conditional Posterior Distribution for Each Parameter
\fontsize{9pt}{8pt}\selectfont
- $\boldsymbol\beta_i\mid (\boldsymbol{\mu}^\top,\boldsymbol{\Sigma},\boldsymbol\gamma^\top,\sigma,\boldsymbol{Y}^\top)$:  
a multivariate normal distribution with mean vector
$$\left(\boldsymbol{\Sigma}^{-1}+\frac{1}{\sigma^2}\sum_{j=1}^{m_i}\mathbf Z_{i,j-1}\mathbf Z_{i,j-1}^\top\right)^{-1}\left(\boldsymbol{\Sigma}^{-1}\boldsymbol\mu+\frac{1}{\sigma^2}\sum_{j=1}^{m_i}(Y_{i,j}-\mathbf{X}_i^\top\boldsymbol\gamma)\mathbf Z_{i,j-1}\right)$$
and covariance matrix $$\boldsymbol{\Sigma}^{-1}+\frac{1}{\sigma^2}\sum_{j=1}^{m_i}\mathbf Z_{i,j-1}\mathbf Z_{i,j-1}^\top$$
- $\boldsymbol{\mu}\mid (\mathbf{B}^\top,\boldsymbol{\Sigma},\boldsymbol\gamma^\top,\sigma,\boldsymbol{Y}^\top)$:  
a multivariate normal distribution with mean vector $$\boldsymbol V(n\boldsymbol V+\boldsymbol\Sigma)^{-1}\left(\sum_{i=1}^n\boldsymbol\beta_i\right)$$ and covariance matrix $$\boldsymbol V(n\boldsymbol V+\boldsymbol\Sigma)^{-1}$$

## Task 2 - Conditional Posterior Distribution for Each Parameter
\fontsize{9pt}{8pt}\selectfont
- $\boldsymbol{\Sigma}\mid (\mathbf{B}^\top,\boldsymbol{\mu}^\top,\boldsymbol\gamma^\top,\sigma,\boldsymbol{Y})$:  
an inverse-Wishart distribution with d.f. $(n+\nu)$ and scale matrix  $$\boldsymbol{S}+\sum_{i=1}^n(\boldsymbol\beta_i-\boldsymbol\mu)(\boldsymbol\beta_i-\boldsymbol\mu)^\top$$
- $\boldsymbol\gamma\mid (\mathbf{B}^\top,\boldsymbol{\mu}^\top,\boldsymbol{\Sigma},\sigma,\boldsymbol{Y}^\top)$:  
a multivariate normal distribution with mean vector $$\left(400\boldsymbol I+\frac{1}{\sigma^2}\sum_{i=1}^nm_i\mathbf{X}_i\mathbf{X}_i^\top\right)^{-1}\left(-\frac{1}{\sigma^2}\sum_{i=1}^n\sum_{j=1}^{m_i}(\mathbf Z_{i,j-1}^\top\boldsymbol\beta_i-Y_{i,j})\mathbf{X}_i\right)$$ and covariance matrix $$\left(400\boldsymbol I+\frac{1}{\sigma^2}\sum_{i=1}^nm_i\mathbf{X}_i\mathbf{X}_i^\top\right)^{-1}$$

## Task 2 - Conditional Posterior Distribution for Each Parameter
\fontsize{9pt}{8pt}\selectfont
- $\sigma\mid \mathbf{B}^\top,\boldsymbol{\mu}^\top,\boldsymbol{\Sigma},\boldsymbol\gamma^\top,\boldsymbol{Y}^\top)$:  
$$\begin{aligned}&\ \pi(\sigma\mid \mathbf{B}^\top,\boldsymbol{\mu}^\top,\boldsymbol{\Sigma},\boldsymbol\gamma^\top,\boldsymbol{Y}^\top)\\
\propto&\ I(\sigma>0)\frac{\sigma^{-\sum_{i=1}^n m_i}}{1+(\sigma/10)^2}\exp\left(-\frac{1}{2\sigma^2}\sum_{i=1}^n\sum_{j=1}^{m_i}(Y_{i,j}-\mathbf Z_{i,j-1}^\top\boldsymbol\beta_i-\mathbf{X}_i^\top\boldsymbol\gamma)^2\right).\end{aligned}$$

## Task 2 - MCMC Algorithm
\begin{algorithm}[H]
  \fontsize{8pt}{6pt}\selectfont
 \caption{MCMC Algorithm}
 \begin{algorithmic}
   \Require $\mathbf{Y}$; $\boldsymbol{\beta}_0, \boldsymbol{\mu}_0, \boldsymbol{\Sigma}_0, \sigma_0, \gamma_0$
   \For{$k=1$ to $10000$}
     \State Gibbs sampling for $\boldsymbol{\beta}_i$'s\par generate ${\boldsymbol{\beta}_i^{(k)}}$ from $\pi(\boldsymbol\beta_i\mid \boldsymbol{\mu}^{(k-1)},\boldsymbol{\Sigma}^{(k-1)},\boldsymbol\gamma^{(k-1)},\sigma^{(k-1)},\boldsymbol{Y}^\top)$
     \State Gibbs sampling for $\boldsymbol{\mu}$\par generate $\boldsymbol{\mu}^{(k)}$ from $\pi(\boldsymbol{\mu}|\mathbf B^{(k)}, \boldsymbol{\Sigma}^{(k-1)},\boldsymbol{\gamma}^{(k-1)}, \sigma^{(k-1)}, \boldsymbol{Y}^\top)$
     \State Gibbs sampling for $\boldsymbol\Sigma$\par generate $\boldsymbol{\Sigma}^{(k)}$ from $\pi(\boldsymbol{\Sigma}| \mathbf B^{(k)}, \boldsymbol{\mu}^{(k)}, \boldsymbol\gamma^{(k-1)}, \sigma^{(k-1)}, \mathbf{Y}^\top)$
     \State Gibbs sampling for $\boldsymbol{\gamma}$\par generate $\boldsymbol{\gamma}^{(k)}$ from $\pi(\boldsymbol{\gamma}| \mathbf B^{(k)}, \boldsymbol{\mu}^{(k)}, \boldsymbol\Sigma^{(k)},\sigma^{(k-1)}, \mathbf{Y}^\top)$
     \State Metropolis-Hastings algorithm for $\sigma$
      \State Propose a conditional distribution of new value $\sigma^*$ from $\text{Uniform}[\sigma^{(k-1)}-a,\sigma^{(k-1)}+a]$, where $a$ is the proposed step
      \State Compute the acceptance ratio\par
      $\lambda=\frac{\pi(\sigma^*|\mathbf{Y}, \boldsymbol{\beta}^{(k)}, \boldsymbol{\mu}^{(k)}, \boldsymbol{\Sigma}^{(k)},\boldsymbol\gamma^{(k)})}{\pi(\sigma^{(k-1)}|\mathbf{Y}, \boldsymbol{\beta}^{(k)}, \boldsymbol{\mu}^{(k)}, \boldsymbol{\Sigma}^{(k)}, \boldsymbol\gamma^{(k)})}$\par
      $\alpha=\min(1,\lambda)$\par
      \State Generate a random number $u$ from $\text{Uniform}(0,1)$
      \State If $u \leq \alpha$, set $\sigma^{(k)} = \sigma^*$, otherwise set $\sigma^{(k)} = \sigma^{(k-1)}$.
    \EndFor
  \State 
  \end{algorithmic}
\end{algorithm}


## Task 2 - Starting Values
\tiny
 \begin{columns}
  \begin{column}{0.4\textwidth}
    \textbf{Final Initial Value Selection and Core Information of MH Algorithm}
    \begin{itemize}
    \item$\boldsymbol{\beta}_i$: This can be obtained through the random effects term in the lmm model. The random effects term can be added to the fixed effects term to obtain $\boldsymbol{\beta}_i^{(0)}$.
    \item$\boldsymbol{\mu}$: This can be obtained through the fixed effects of intercept, windpre, latdiff, longdiff, winddiff term.
    \item$\boldsymbol{\gamma}$: This can be obtained through the fixed effects of Year, Active Month, and Nature term.
    \item$\sigma$: This can be obtained through the model residual.
    \item$\boldsymbol\Sigma$: This can be obtained through the `VarCorr` function which returns the covariance matrix of the random effects in the model $\boldsymbol\Sigma^{^{(0)}}$.
    \end{itemize}
  \end{column}
 \begin{column}{0.6\textwidth}
 \centering
   \begin{table}[h]
   \centering
   \caption{Initial Value Setting}
   \begin{tabular}{|c|c|}
   \hline
   Parameter & Value \\
   \hline
   $\boldsymbol{\mu}^\top$ & $(24.25, 0.94, -0.02, -0.24, 0.47)$ \\
   $\boldsymbol{\gamma}^\top$ & $(-0.01, 0.35, 0.28, 0.37, 0.12, 0.08)$ \\
   $\boldsymbol\Sigma$ & $\begin{pmatrix}
   0.358 & -0.010 & 0.039 & 0.121 & 0.028 \\
   -0.01 & 0.001 & -0.003 & -0.005 & 0.002 \\
   0.039 & -0.003 & 0.043 & 0.034 & -0.019 \\
   0.121 & -0.005 & 0.034 & 0.069 & 0.003 \\
   0.028 & 0.002 & -0.019 & 0.003 & 0.017 \\
    \end{pmatrix}$ \\
   $\sigma$ & 5.27 \\
   \hline
   \end{tabular}
   \end{table}
  \end{column}
 \end{columns}
 
## Task 2 - MCMC Algorithm R code
```{r eval=FALSE, size= "tiny"}
# Gibbs sampling
B_sample <- function(mu, Sigma, gamma, sigma) {
  Sigma.inv <- solve(Sigma)
  B_mean_cov <- function(i) {
    cov <- solve(Sigma.inv + 1/sigma^2 * t(Z[[i]]) %*% Z[[i]])
    mean <- cov %*% (Sigma.inv %*% mu + 1/sigma^2 * colSums((Y[[i]] - (X[i,] %*% gamma)[,]) * Z[[i]]))
    list(mean = mean, cov = cov)
  }
  mean_cov_list <- lapply(1:n, B_mean_cov)
  B <- sapply(mean_cov_list, function(x) {mvrnorm(mu = x$mean, Sigma = x$cov)})
  return(B)
}

# MH algorithm (random walk)
sigma_sample <- function(sigma, B, gamma, a) {
  sigma_new <- sigma + (runif(1) - 0.5) * 2 * a # candidate sigma
  if (sigma_new <= 0) {
    return(sigma)
  }
  RSS <- sum(sapply(1:n, function(i) sum((Y[[i]] - Z[[i]] %*% B[,i] - (X[i,] %*% gamma)[,])^2)))
  log_kernal_ratio <- -sum(m) * log(sigma_new/sigma) +
    log(1 + (sigma/10)^2) - log(1 + (sigma_new/10)^2) -
    0.5 * (1/sigma_new^2 - 1/sigma^2) * RSS
  log_prob <- min(0, log_kernal_ratio)
  sigma <- ifelse(log_prob > log(runif(1)), sigma_new, sigma)
  return(sigma)
}
```

## Task 2 - Results Presentation (hyperparameters)
- search window $a=0.1$
- burn-in $=8000$
- resulting acceptance rate $=0.418$

## Task 2 - Results Presentation (Parameters (Burn-In 8000))
\begin{figure} 
  \centering 
  \subfigure[mu]{ 
    \label{sub1}
    \includegraphics[width=2.0in, height = 1.4in]{parameters_burnin/Mu_parameters_time_series_plot.png} 
  } 
  \subfigure[Sigma matrix]{ 
    \label{sub2} 
    \includegraphics[width=2.0in, height = 1.4in]{parameters_burnin/Sigma_parameters_time_series_plot.png} 
  } 
  \subfigure[gamma]{ 
    \label{sub3} 
    \includegraphics[width=2.0in, height = 1.4in]{parameters_burnin/Gamma_parameters_time_series_plot.png} 
  } 
  \subfigure[sigma]{ 
    \label{sub4} 
    \includegraphics[width=2.0in, height = 1.4in]{parameters_burnin/igma_parameters_time_series_plot.png} 
  } 
  \label{para1} 
\end{figure}

## Task 2 - Results Presentation (Histogram Plots for $\boldsymbol \mu$)
```{r warning=FALSE, echo=FALSE}
mu_list <- read.csv("./data/mu_list.csv")
# Get the last 2000 rows of the mu_list data
mu_data <- tail(mu_list, 2000)

# Convert the mu_data to a data frame
mu_df <- as.data.frame(mu_data)

# Set the number of bins for the histogram
num_bins <- 30

# Define a color palette for the histograms
colors <- c("red", "green", "blue", "orange", "purple")

# Plot the histograms for each column in mu_data using ggplot
histograms <- lapply(seq_along(mu_df), function(i) {
  ggplot(mu_df, aes(x = mu_df[, i])) +
    geom_histogram(bins = num_bins, color = "black", fill = colors[i], alpha = 0.7) +
    ggtitle(paste0("mu_", i - 1)) +
    xlab("Value")
})

# Combine the histograms into a single plot using gridExtra
grid.arrange(grobs = histograms, ncol = 3)
```

##  Task 2 - Results Presentation (Parameter Estimations of $\boldsymbol \mu$ and $\boldsymbol \Sigma$)
\centering
\tiny
\textbf{$\hat{\boldsymbol\mu}$}
\[
\begin{pmatrix}
  \hat{\mu}_0 & -3.6594881 \\
  \hat{\mu}_1 & 0.9020437 \\
  \hat{\mu}_2 & -0.0584559 \\
  \hat{\mu}_3 & -0.4502597 \\
  \hat{\mu}_4 & 0.4714633
\end{pmatrix}
\]

\centering
\tiny
\textbf{$\hat{\boldsymbol\Sigma}$}
\[
\begin{pmatrix}
  & 0.7521 & -0.0155 & -0.0860 & 0.0136 & -0.0060\\
  & -0.0155 & 0.0050 & -0.0023 & -0.0013 & 0.0006\\
  & -0.0860 & -0.0023 & 0.2705 & -0.0085 & -0.0024\\
  & 0.0136 & -0.0013 & -0.0085 & 0.1287 & 0.0057\\
  & -0.0060 & 0.0006 & -0.0024 & 0.0057 & 0.0268
\end{pmatrix}
\]


##  Task 2 - Results Presentation (CIs of $\boldsymbol \mu$)
```{r warning=FALSE, echo=FALSE}
data <- data.frame(
  variable = c("mu_0", "mu_1", "mu_2", "mu_3", "mu_4"),
  mean = c(-3.6594881, 0.9020437, -0.0584559, -0.4502597, 0.4714633),
  lower_CI = c(-4.9823968, 0.8379360, -0.2942283, -0.5789892, 0.3958291),
  upper_CI = c(-1.1458245, 0.9670541, 0.1610285, -0.3093745, 0.5443864)
)
knitr::kable(data, format = "markdown")

```

## Task 3
**Objective:**

Compute posterior summaries and 95% credible intervals of $\gamma$, the fixed effects associated with the covariates in the model. Using the estimated Bayesian model, answer the following questions:\par
(1) Are there seasonal differences in hurricane wind speeds?\par
(2) Is there evidence to support the claim that hurricane wind speeds have been increasing over the years?

## Task 3 - Parameters Convergence
\begin{figure}[H] 
\includegraphics[width=1.0\textwidth]{jacky_part_pic/gamma.png} 
\end{figure}

## Task 3 - 95% Credible Intervals
\begin{figure}[H] 
\includegraphics[width=1.0\textwidth]{jacky_part_pic/95ci.png} 
\end{figure}
- Conclusion: 

  1. There is no seasonal difference (active v.s inactive) in hurricane wind speeds.
  
  2. There is no evidence to support the claim that hurricane wind speeds have been increasing over the years because of divergence.
  
## Task 4 - Objective
**Objective:**

With the estimated model parameters and covariate values, you can calculate the predicted wind speed for each time point using the model equation. This way, you can track the hurricane and compare the predicted wind speeds with the actual wind speeds recorded during the hurricane. Please evaluate how well the estimated Bayesian model can track individual hurricanes.

## Task 4
**Prediction:**
Using the parameters after burn-in, we can obtain the predicted value for each hurricane.
$$\hat Y_{i}(t+6) =\hat \beta_{0,i}+\hat \beta_{1,i}Y_{i}(t) + \hat \beta_{2,i}\Delta_{i,1}(t)+
\hat \beta_{3,i}\Delta_{i,2}(t) +\hat \beta_{4,i}\Delta_{i,3}(t)  + \mathbf{X}_i {^\top \hat {\boldsymbol\gamma}}$$ 

**Performance evaluation:**
For each hurricane, we can evaluate the estimated Bayesian model performance by calculating
$$
RMSE = \sqrt{\frac{\sum_{i=1}^{n} (y_i - \hat{y_i})^2}{n}}
$$
$$
R^2 = 1 - \frac{\sum_{i=1}^{n} (y_i - \hat{y_i})^2}{\sum_{i=1}^{n} (y_i - \bar{y})^2}
$$

## Task 4
\begin{table}[h]
\centering
\small
\caption{Summary of RMSE and R-squared for selected hurricanes}
\label{table:summary}
\begin{tabular}{llcc}
\hline
\textbf{ID} & \textbf{Year} & \textbf{RMSE} & \textbf{R-squared} \\ \hline
ABBY.1960 & 1960 & 8.8804 & 0.7700 \\
ABBY.1964 & 1964 & 9.6430 & 0.3033 \\
ABBY.1968 & 1968 & 3.5043 & 0.9360 \\
ABLE.1950 & 1950 & 3.6755 & 0.9813 \\
ABLE.1951 & 1951 & 3.4802 & 0.9767 \\
ABLE.1952 & 1952 & 4.5183 & 0.9583 \\
AGNES.1972 & 1972 & 5.2483 & 0.8881 \\
ALBERTO.1982 & 1982 & 8.0473 & 0.7499 \\
ALBERTO.1988 & 1988 & 2.6121 & 0.7420 \\
ALBERTO.1994 & 1994 & 4.3941 & 0.8807 \\
ALBERTO.2000 & 2000 & 3.7896 & 0.9625 \\
ALBERTO.2006 & 2006 & 4.3591 & 0.7882 \\
ALBERTO.2012 & 2012 & 3.2193 & 0.8036 \\
ALEX.1998 & 1998 & 2.9351 & 0.7289 \\
ALEX.2004 & 2004 & 5.4552 & 0.9539 \\ \hline
\end{tabular}
\end{table}

## Task 4
Prediction performance on random chosen example hurricanes.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{predic_plots/b_merge.png}
  \caption{Time series prediction plot}
\end{figure}

## Task 4
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{predic_plots/a_1_4.png}
  \caption{Prediction vs. observation}
\end{figure}

## Task 4
**Performance evaluation:**
We plot the $RMSE$ and $R^2$ distribution for all the hurricanes
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{predic_plots/rmse1.png}
  \caption{RMSE distribution}
\end{figure}

## Task 4
**Performance evaluation:**
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{predic_plots/R_square.png}
  \caption{$R^2$ distribution}
\end{figure}

## Task 4
To enhance the evaluation of our prediction performance, we examined the distribution of RMSE values across various properties of hurricanes.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{predic_plots/rmse_nature.png}
  \caption{RMSE under different natures}
\end{figure}

## Task 4
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{predic_plots/nature2.png}
  \caption{RMSE under different natures}
\end{figure}

## Task 4
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{predic_plots/months1.png}
  \caption{RMSE under different months}
\end{figure}

## Task 4
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.8\textwidth]{predic_plots/month2.png}
  \caption{RMSE under different months}
\end{figure}


## Discussions

- Better parameters convergence performance in MCMC

- Prediction latency in responding to wind speed change

## Q&A

- Thanks for listening!