---
title: "MCMC"
output: 
    pdf_document
---

```{r include=FALSE, echo=FALSE}
library(MASS)
library(tidyverse)
library(nlme)
library(lme4)
library(ggplot2)
library(gridExtra)
library(cowplot)
library(knitr)
```

### Data preprocessing (same as EDA)

```{r}
origin_df <- read.csv("hurrican703.csv")

hurricane_df <- origin_df %>% 
  mutate(
    Active = ifelse(Month %in% month.name[8:10], "Active", "Inactive"),
    Active = factor(Active, levels = c("Inactive", "Active")),
    Month = factor(Month, levels = month.name[-c(2:3)]), # April-January (January ref, may choose another)
    Nature = as.factor(Nature), # TS,ET,DS,SS,NR (DS ref, may choose another)
    # note: one hurricane can have multiple natures throughout its life
    time = gsub("[()]", "", time),
    time = paste0(ifelse(substr(time, 1, 2) > 23, "19", "20"), time),
    time = as.POSIXct(time, format = "%Y-%m-%d %H:%M:%S"),
    hour = substr(time, 12, 19)
  ) %>% 
  # remove data not at six-hour time intervals. (613 observations)
  filter(hour %in% c("00:00:00", "06:00:00", "12:00:00", "18:00:00")) %>% 
  dplyr::select(-hour)

# remove hurricanes that has only 2 (<3) observations (change the threshold if you wish)
few_id <- hurricane_df %>% 
  group_by(ID) %>% 
  summarize(obs = n()) %>% 
  filter(obs < 3) %>% 
  .$ID
hurricane_df <- hurricane_df %>% filter(!(ID %in% few_id)) # remove 3 hurricanes

# manually correct hurricanes that have same names but are actually different
hurricane_df <- 
  hurricane_df %>% 
  mutate(
    # 2 hurricanes with the name ALICE.1954
    ID = ifelse(ID == "ALICE.1954" & Month == "June", "ALICE.1954(1)", ID),
    ID = ifelse(ID == "ALICE.1954", "ALICE.1954(2)", ID),
    # 4 hurricanes with the name SUBTROP:UNNAMED.1974
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974" & Month == "June", "SUBTROP:UNNAMED.1974(1)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974" & Month == "July", "SUBTROP:UNNAMED.1974(2)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974" & Month == "August", "SUBTROP:UNNAMED.1974(3)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974", "SUBTROP:UNNAMED.1974(4)", ID),
    # 2 hurricanes with the name SUBTROP:UNNAMED.1976
    ID = ifelse(ID == "SUBTROP:UNNAMED.1976" & Month == "May", "SUBTROP:UNNAMED.1976(1)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1976", "SUBTROP:UNNAMED.1976(2)", ID)
  )
```

### Data partition (may be useless.)

```{r}
# data partition, no need to partition data
Training <- hurricane_df
```

### Create $\mathbf{X}$, $\mathbf{Y}$, $\mathbf{Z}$ and $\boldsymbol{m}=(m_i)$ in R

```{r}
n <- length(unique(Training$ID))

Y <- split(Training$Wind.kt, Training$ID) %>% 
  lapply(function(x) x[-c(1:2)])

X <- Training %>% 
  group_by(ID) %>% 
  slice(1) %>% 
  dplyr::select(ID, Season, Active, Nature) %>% 
  ungroup(ID)
X <- model.matrix(~., X[-1])[,-1]

Z <- Training %>% 
  group_by(ID) %>% 
  mutate(
    intercept = 1,
    wind_pre = lag(Wind.kt),
    lat_diff = lag(Latitude) - lag(Latitude, 2),
    long_diff = lag(Longitude) - lag(Longitude, 2),
    wind_diff = lag(Wind.kt) - lag(Wind.kt, 2),
  ) %>% 
  drop_na %>% 
  dplyr::select(ID, intercept, wind_pre, lat_diff, long_diff, wind_diff)
Z <- split(Z[, names(Z)[-1]], Z$ID) %>% 
  lapply(as.matrix)

m <- Training %>% 
  group_by(ID) %>% 
  summarize(obs = n() - 2) %>% 
  .$obs
```

### MCMC

```{r}
B_sample <- function(mu, Sigma, gamma, sigma) {
  Sigma.inv <- solve(Sigma)
  B_mean_cov <- function(i) {
    cov <- solve(Sigma.inv + 1/sigma^2 * t(Z[[i]]) %*% Z[[i]])
    mean <- cov %*% (Sigma.inv %*% mu + 1/sigma^2 * colSums((Y[[i]] - (X[i,] %*% gamma)[,]) * Z[[i]]))
    list(mean = mean, cov = cov)
  }
  mean_cov_list <- lapply(1:n, B_mean_cov)
  B <- sapply(mean_cov_list, function(x) {mvrnorm(mu = x$mean, Sigma = x$cov)})
  return(B)
}

mu_sample <- function(B, Sigma) {
  cov <- V %*% solve(n*V + Sigma)
  mean <- cov %*% rowSums(B)
  mu <- mvrnorm(mu = mean, Sigma = cov)
  return(mu)
}

Sigma_sample <- function(B, mu) {
  Sigma.inv <- rWishart(n = 1, Sigma = solve(S + (B - mu) %*% t(B - mu)), df = n + nu)[,,]
  Sigma <- solve(Sigma.inv)
  return(Sigma)
}

gamma_sample <- function(B, sigma) {
  X_trans <- sqrt(m) * X
  cov <- solve(400*diag(6) + 1/sigma^2 * t(X_trans) %*% X_trans)
  total <- rowSums(sapply(1:n, function(i) sum(Z[[i]] %*% B[,i] - Y[[i]]) * X[i,]))
  mean <- cov %*% (-1/sigma^2 * total)
  gamma <- mvrnorm(mu = mean, Sigma = cov)
  return(gamma)
}

# MH algorithm (random walk)
sigma_sample <- function(sigma, B, gamma, a) {
  sigma_new <- sigma + (runif(1) - 0.5) * 2 * a # candidate sigma
  if (sigma_new <= 0) {
    return(sigma)
  }
  RSS <- sum(sapply(1:n, function(i) sum((Y[[i]] - Z[[i]] %*% B[,i] - (X[i,] %*% gamma)[,])^2)))
  log_kernal_ratio <- -sum(m) * log(sigma_new/sigma) +
    log(1 + (sigma/10)^2) - log(1 + (sigma_new/10)^2) -
    0.5 * (1/sigma_new^2 - 1/sigma^2) * RSS
  log_prob <- min(0, log_kernal_ratio)
  sigma <- ifelse(log_prob > log(runif(1)), sigma_new, sigma)
  return(sigma)
}

MCMC <- function(B0, mu0, Sigma0, gamma0, sigma0, a, iter) {
  B <- B0
  mu <- mu0
  Sigma <- Sigma0
  gamma <- gamma0
  sigma <- sigma0
  res <- vector("list", iter)
  for (i in 1:iter) {
    B <- B_sample(mu, Sigma, gamma, sigma)
    mu <- mu_sample(B, Sigma)
    Sigma <- Sigma_sample(B, mu)
    gamma <- gamma_sample(B, sigma)
    sigma <- sigma_sample(sigma, B, gamma, a)
    res[[i]] <- list(B = B, mu = mu, Sigma = Sigma, gamma = gamma, sigma = sigma)
  }
  return(res)
}
```

### Fit lmm model to select starting values

```{r}
df <- hurricane_df %>% 
  group_by(ID) %>% 
  mutate(
    Season = first(Season),
    Active = first(Active),
    Nature = first(Nature),
    wind_pre = lag(Wind.kt),
    lat_diff = lag(Latitude) - lag(Latitude, 2),
    long_diff = lag(Longitude) - lag(Longitude, 2),
    wind_diff = lag(Wind.kt) - lag(Wind.kt, 2),
  ) %>% 
  drop_na %>% 
  dplyr::select(ID, Wind.kt, Season, Active, Nature, wind_pre, lat_diff, long_diff, wind_diff)

lmm <- lmer(Wind.kt ~ Season + Active + Nature + wind_pre + lat_diff + long_diff + wind_diff +
            (1 + wind_pre + lat_diff + long_diff + wind_diff | ID),
            data = df)
summary(lmm)

mu0 <- as.vector(fixed.effects(lmm)[-c(2:7)])
gamma0 <- as.vector(fixed.effects(lmm)[2:7])
n <- length(unique(df$ID))
B0 <- matrix(rep(mu0, n), nrow = 5)
B0 <- B0 + t(as.matrix(random.effects(lmm)$ID))
sigma0 <- sd(residuals(lmm))
Sigma0 <- as.matrix(Matrix::bdiag(VarCorr(lmm))) + diag(5)*0.001 # computationally singular
```


### Apply MCMC on training data

1. Please select appropriate $\boldsymbol V$ and $\boldsymbol S$ values.
2. Please select appropriate starting values for parameters, especially $\mathbf{B}$.
3. You can save the results `xxx_list` as csv files or otherwise. Note that we need to find the answers to the above 2 problems and also the following questions to make sure that we won't make any updates for these saved data:

* Should we apply MCMC on training hurricanes or all hurricanes? (in other words, should we partition the data into training and test data?) This relies on whether we hope to make predictions only on test hurricanes or on hurricanes used for MCMC.
* Given that samples of $\mathbf{B}$ is a 10000*2815 matrix, and we only use $\mathbf{B}$ for predictions on hurricanes used for MCMC, we can decide which hurricanes we are gonna predict (e.g., hurricanes having appropriate number of observations, famous hurricanes, hurricanes of different natures, years, months), and only save the samples of corresponding $\boldsymbol\beta_i$'s.
* We need to run the MCMC function and save the results (`xxx_list`) after we solve the above 3 problems. Then we need to check whether the `a` in the random walk for sampling $\sigma$ is appropriate. (acceptance rate 30% - 60%, see far below) If not, we need to change `a` and regenerate the results.

estimated running time: 10 mins for 10,000 iterations


### Fit lmm model to select starting values

```{r}
# constants
nu <- 5 # suggested
V <- diag(5) # can try other. see requirement
S <- diag(5) # can try other. see requirement
```

```{r}
set.seed(1)
res <- MCMC(B0, mu0, Sigma0, gamma0, sigma0, a = 0.1, iter = 10000) # try larger values

# B. beta_i_j: i: ith hurricane, j: 1-5
B_list <- t(mapply(function(x) x$B, res))
B_names <- apply(expand.grid(0:4, 1:n), 1,
                 function(x) paste("beta", x[2], x[1], sep = "_"))
colnames(B_list) <- B_names
# mu
mu_list <- t(mapply(function(x) x$mu, res))
colnames(mu_list) <- colnames(Z[[1]])
# Sigma (symmetric)
Sigma_list <- t(mapply(function(x) x$Sigma, res))
Sigma_names <- apply(expand.grid(1:5, 1:5), 1,
                     function(x) paste("Sigma_", x[2], x[1], sep = ""))
colnames(Sigma_list) <- Sigma_names
# gamma (month ref: January, nature ref: DS)
gamma_list <- t(mapply(function(x) x$gamma, res))
# sigma
sigma_list <- mapply(function(x) x$sigma, res)


B_mean = bind_rows(colSums(B_list[5001:10000,])/5000, colSums(B_list[6001:10000,])/4000,
                  colSums(B_list[7001:10000,])/3000, colSums(B_list[8001:10000,])/2000,
                  colSums(B_list[9001:10000,])/1000)



write.csv(B_list, file = "./data/B_list.csv", row.names = FALSE) # last 5000 obs
write.csv(B_mean, file = "./data/B_list_lastmean.csv", row.names = FALSE) # last 5000 obs
write.csv(mu_list, file = "./data/mu_list.csv", row.names = FALSE)
write.csv(Sigma_list, file = "./data/Sigma_matrix_list.csv", row.names = FALSE)
write.csv(gamma_list, file = "./data/gamma_list.csv", row.names = FALSE)
write.csv(sigma_list, file = "./data/sigma_list.csv", row.names = FALSE)
```

### Resulting plots

Please make time series plots for all parameters to see if and from which index their samples converge. (may not need for $\mathbf{B}$, since it has $5n$ parameters (5 for each hurricane))

Please also make autocorrelation plots.

```{r fig.height=5, fig.width=8}
beta_list <- read.csv("./data/B_list_lastmean.csv")
mu_list <- read.csv("./data/mu_list.csv")
Sigma_list <- read.csv("./data/Sigma_matrix_list.csv")
gamma_list <- read.csv("./data/gamma_list.csv")
sigma_list <- read.csv("./data/sigma_list.csv")$x
```

#### Select burning time
```{r}
n_burnin=5000
```


### trace plot
#### Mu
```{r}
# Create the folder
dir.create(file.path("parameters_burnin", "mu"), showWarnings = FALSE)

# Create a list to store the plots
plots <- list()

# Loop through the columns of mu_list
for (i in 1:ncol(mu_list)) {
  # Create a plot and add it to the list
  p <- ggplot(data = data.frame(y = mu_list[1:nrow(Sigma_list), i])) +
    geom_line(aes(x = 1:length(y), y = y)) +
    xlab("") +
    ylab(paste0("Mu ", i)) +
    theme_bw()+
    geom_vline(xintercept = n_burnin, linetype = "dashed", color = "red")
   plots[[i]] <- p
  
  # Save the plot as a PNG file
  filename <- file.path("parameters_burnin", "mu", paste0("parameter_", i, ".png"))
  ggsave(filename, plot = p, width = 8, height = 6)
}

# Merge the plots into a single plot
merged_plot <- plot_grid(plotlist = plots, ncol = 1)

# Add a title to the merged plot
title <- ggdraw() + draw_label("Time Series Plot of Mu", fontface = "bold", size = 14) + theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

# Combine the title and the plot
final_plot <- plot_grid(title, merged_plot, ncol = 1, align = "v", axis = "tb", rel_heights = c(0.1, 0.9))

# Save the merged plot as a PNG file
ggsave(file.path("parameters_burnin", "Mu_parameters_time_series_plot.png"), plot = final_plot, width = 12, height = 8)
```

# Sigma
```{r}
# Create the folder
dir.create(file.path("parameters_burnin", "Sigma"), showWarnings = FALSE)

# Create a list to store the plots
plots <- list()

# Loop through the columns of Sigma_list
for (i in 1:ncol(Sigma_list)) {
  # Create a plot and add it to the list
  p <- ggplot(data = data.frame(y = Sigma_list[1:nrow(Sigma_list), i])) +
    geom_line(aes(x = 1:length(y), y = y)) +
    xlab("") +
    ylab(paste0("Sigma ", i)) +
    theme_bw()+
    geom_vline(xintercept = n_burnin, linetype = "dashed", color = "red")
  plots[[i]] <- p
  
  # Save the plot as a PNG file
  filename <- file.path("parameters_burnin", "Sigma", paste0("parameter_", i, ".png"))
  ggsave(filename, plot = p, width = 8, height = 6)
}

# Merge the plots into a single plot
merged_plot <- plot_grid(plotlist = plots, ncol = 5)

# Add a title to the merged plot
title <- ggdraw() + draw_label("Time Series Plot of Sigma", fontface = "bold", size = 14) + theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

# Combine the title and the plot
final_plot <- plot_grid(title, merged_plot, ncol = 1, align = "v", axis = "tb", rel_heights = c(0.1, 0.9))

# Save the merged plot as a PNG file
ggsave(file.path("parameters_burnin", "Sigma_parameters_time_series_plot.png"), plot = final_plot, width = 12, height = 8)
```


#### Gamma

```{r}
# Create the folder
dir.create(file.path("parameters_burnin", "gamma gride"), showWarnings = FALSE)

# Create a list to store the plots
plots <- list()

# Loop through the columns of gamma_list
for (i in 1:ncol(gamma_list)) {
  # Create a plot and add it to the list
  p <- ggplot(data = data.frame(y = gamma_list[1:nrow(Sigma_list), i])) +
    geom_line(aes(x = 1:length(y), y = y)) +
    xlab("") +
    ylab(paste0("Gamma", i)) +
    theme_bw()+
    geom_vline(xintercept = n_burnin, linetype = "dashed", color = "red")
  plots[[i]] <- p
  # Save the plot as a PNG file
  filename <- file.path("parameters_burnin", "gamma", paste0("parameter_", i, ".png"))
  ggsave(filename, plot = p, width = 8, height = 6)
}

# Merge the plots into a single plot
merged_plot <- grid.arrange(grobs = plots, ncol = 2)

# Add a title to the merged plot
title <- ggdraw() + draw_label("Time Series Plot of Gamma", fontface = "bold", size = 14) + theme(plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm"))

# Combine the title and the plot
final_plot <- plot_grid(title, merged_plot, ncol = 1, align = "v", axis = "tb", rel_heights = c(0.1, 0.9))

# Save the merged plot as a PNG file
ggsave(file.path("parameters_burnin", "Gamma_parameters_time_series_plot.png"), plot = final_plot, width = 12, height = 8)

```


#### Visualization for sigma
```{r}
ggplot(data = data.frame(sigma = sigma_list[1:length(sigma_list)]), aes(x = 1:length(sigma), y = sigma)) +
  geom_line() +
  xlab("") +
  ylab("sigma") +
  theme_bw() +
  geom_vline(xintercept = n_burnin, linetype = "dashed", color = "red")
ggsave(file.path("parameters_burnin","igma_parameters_time_series_plot.png"), width = 8, height = 6, dpi = 300)
```

### Histogram
#### Mu
```{r}
mu_data <- tail(mu_list, 10000-n_burnin)

# Convert the mu_data to a data frame
mu_df <- as.data.frame(mu_data)

mu_data
# Set the number of bins for the histogram
num_bins <- 30

# Define a color palette for the histograms
colors <- c("red", "green", "blue", "orange", "purple")

# Plot the histograms for each column in mu_data using ggplot
histograms <- lapply(seq_along(mu_df), function(i) {
  ggplot(mu_df, aes(x = mu_df[, i])) +
    geom_histogram(bins = num_bins, color = "black", fill = colors[i], alpha = 0.7) +
    ggtitle(paste0("mu_", i - 1)) +
    xlab("Value")
})

# Combine the histograms into a single plot using gridExtra
combined_plot <- grid.arrange(grobs = histograms, ncol = 3)
file_path <- file.path("parameters_burnin", "Mu_Parameters_Histogram_Plot.png")
# Save the plot as a PNG file
ggsave(file_path, plot = combined_plot, width = 11, height = 8.5, units = "in")


```


```{r}
sigma_data <- tail(Sigma_list, 10000-n_burnin)

# Convert the sigma_data to a data frame
sigma_df <- as.data.frame(sigma_data)

# Set the number of bins for the histogram
num_bins <- 30

# Define a color palette for the histograms
colors <- c("red", "green", "blue", "orange", "purple",
            "brown", "gray", "pink", "black", "magenta",
            "cyan", "yellow", "darkred", "darkgreen", "darkblue",
            "darkorange", "blue", "darkgray", "orange", "lightblue",
            "lightgreen", "lightpink", "lightgray", "purple", "darkred")

# Plot the histograms for each column in sigma_data using ggplot
histograms <- lapply(seq_along(sigma_df), function(i) {
  ggplot(sigma_df, aes(x = sigma_df[, i])) +
    geom_histogram(bins = num_bins, color = "black", fill = colors[i], alpha = 0.7) +
    ggtitle(paste0("Sigma_", i - 1)) +
    xlab("Value")
})

# Combine the histograms into a single plot using gridExtra
combined_plot <- grid.arrange(grobs = histograms, ncol = 5)
file_path <- file.path("parameters_burnin", "Sigma_Parameters_Histogram_Plot.png")
# Save the plot as a PNG file
ggsave(file_path, plot = combined_plot, width = 11, height = 8.5, units = "in")
```


### Auto correlationplot

```{r}
# Set the lag.max for the acf function
lag.max <- 2500

# Define a color palette for the plots
colors <- c("red", "green", "blue", "orange", "purple")

# Plot the autocorrelation plot for each column in mu_df using ggplot
autocorrelation_plots <- lapply(seq_along(mu_df), function(i) {
  acf_values <- acf(mu_df[, i], lag.max = lag.max, plot = FALSE)$acf
  ggplot(data.frame(lag = seq(0, lag.max)), aes(x = lag, y = acf_values)) +
    geom_hline(yintercept = 0, color = "black") +
    geom_segment(aes(xend = lag, yend = 0), color = colors[i], alpha = 0.7) +
    ggtitle(paste0("Autocorrelation Plot for mu_", i - 1)) +
    xlab("Lag") +
    ylab("Autocorrelation")
})

# Combine the plots into a single plot using gridExtra
grid.arrange(grobs = autocorrelation_plots, ncol = 3)

```

```{r}
# Set the lag.max for the acf function
lag.max <- 2500

# Define a color palette for the plots
colors <- c("red", "green", "blue", "orange", "purple",
            "brown", "gray", "pink", "black", "magenta",
            "cyan", "yellow", "darkred", "darkgreen", "darkblue",
            "darkorange", "blue", "darkgray", "orange", "lightblue",
            "lightgreen", "lightpink", "lightgray", "purple", "darkred")

# Plot the autocorrelation plot for each column in sigma_df using ggplot
autocorrelation_plots <- lapply(seq_along(sigma_df), function(i) {
  acf_values <- acf(sigma_df[, i], lag.max = lag.max, plot = FALSE)$acf
  ggplot(data.frame(lag = seq(0, lag.max)), aes(x = lag, y = acf_values)) +
    geom_hline(yintercept = 0, color = "black") +
    geom_segment(aes(xend = lag, yend = 0), color = colors[i], alpha = 0.7) +
    ggtitle(paste0("Autocorrelation Plot for Sigma_", i - 1)) +
    xlab("Lag") +
    ylab("Autocorrelation")
})

# Combine the plots into a single plot using gridExtra
combined_plot <- grid.arrange(grobs = autocorrelation_plots, ncol = 5)

# Construct the file path
file_path <- file.path("parameters_burnin", "Sigma_Parameters_ACF_Plot.png")

# Save the plot as a PNG file in the "parameters_burnin" directory
ggsave(file_path, plot = combined_plot, width = 11, height = 8.5, units = "in")


```

```{r}
gamma_data <- tail(gamma_list, 5000)

# Convert the gamma_data to a data frame
gamma_df <- as.data.frame(gamma_data)

# Set the lag.max for the acf function
lag.max <- 2500

# Define a color palette for the plots
colors <- c("red", "green", "blue", "orange", "purple","grey")

# Plot the autocorrelation plot for each column in gamma_df using ggplot
autocorrelation_plots <- lapply(seq_along(gamma_df), function(i) {
  acf_values <- acf(gamma_df[, i], lag.max = lag.max, plot = FALSE)$acf
  ggplot(data.frame(lag = seq(0, lag.max)), aes(x = lag, y = acf_values)) +
    geom_hline(yintercept = 0, color = "black") +
    geom_segment(aes(xend = lag, yend = 0), color = colors[i], alpha = 0.7) +
    ggtitle(paste0("Autocorrelation Plot for gamma_", i - 1)) +
    xlab("Lag") +
    ylab("Autocorrelation")
})

# Combine the plots into a single plot using gridExtra
combined_plot <- grid.arrange(grobs = autocorrelation_plots, ncol = 2)

# Construct the file path
file_path <- file.path("parameters_burnin", "Gamma_Parameters_ACF_Plot.png")

# Save the plot as a PNG file in the "parameters_burnin" directory
ggsave(file_path, plot = combined_plot, width = 11, height = 8.5, units = "in")

```

```{r}
sigma_data <- tail(sigma_list, 10000-n_burnin)

# Extract the single column of data from the sigma_data list
sigma_vec <- sigma_data

# Set the lag.max for the acf function
lag.max <- 2500

# Define a color palette for the plots
colors <- c("red")

# Calculate the autocorrelation values using acf
acf_values <- acf(sigma_vec, lag.max = lag.max, plot = FALSE)$acf

# Plot the autocorrelation function using ggplot
combined_plot <- ggplot(data.frame(lag = seq(0, lag.max)), aes(x = lag, y = acf_values)) +
  geom_hline(yintercept = 0, color = "black") +
  geom_segment(aes(xend = lag, yend = 0), color = colors[1], alpha = 0.7) +
  ggtitle("Autocorrelation Plot for Sigma") +
  xlab("Lag") +
  ylab("Autocorrelation")

# Construct the file path
file_path <- file.path("parameters_burnin", "igma_Parameters_ACF_Plot.png")

# Save the plot as a PNG file in the "parameters_burnin" directory
ggsave(file_path, plot = combined_plot, width = 11, height = 8.5, units = "in")
```

```{r}
# Combine the histograms into a single plot using gridExtra
library(gridExtra)
grid.arrange(grobs = histograms, ncol = 5)

sigma_mean <- apply(sigma_data, 2, mean)
Sigma <- matrix(c(0.7521122437, -0.0154712382, -0.0860070724, 0.0135698462, -0.0060100050,
                  -0.0154712382, 0.0049778296, -0.0023300326, -0.0013199897, 0.0005661384,
                  -0.0860070724, -0.0023300326, 0.2705120482, -0.0085330838, -0.0023624339,
                  0.0135698462, -0.0013199897, -0.0085330838, 0.1287440122, 0.0056527804,
                  -0.0060100050, 0.0005661384, -0.0023624339, 0.0056527804, 0.0268370062),
                nrow = 5, ncol = 5, byrow = TRUE)
sds <- sqrt(diag(Sigma)) 
rho <- Sigma / (sds %*% t(sds)) 

```

```{r warning=FALSE, echo=FALSE}
mu_list <- read.csv("./data/mu_list.csv",skip=5000)
mean_vec_mu <- colMeans(mu_list)
lower_ci_vec_mu <- apply(mu_list, 2, quantile, 0.05)
upper_ci_vec_mu <- apply(mu_list, 2, quantile, 0.95)

output_df_mu <- data.frame(
  Parameter = paste0("mu_", 0:4),
  Mean = mean_vec_mu,
  Lower_CI = lower_ci_vec_mu,
  Upper_CI = upper_ci_vec_mu
)

colnames(output_df_mu)[-1] <- paste0("mu_mean_id", 0:4)

kable(output_df_mu, format="html")

```
