---
title: "Task 4"
output: 
    pdf_document
---

```{r setup, include=FALSE, echo=FALSE}
options(warn = -1, message = FALSE)

library(MASS)
library(tidyverse)
library(ggridges)
library(dplyr)
library(broom)
library(gridExtra)
```

**Task 2:** With the estimated model parameters and covariate values, you can calculate the predicted wind speed for each time point using the model equation. This way, you can track the hurricane and compare the predicted wind speeds with the actual wind speeds recorded during the hurricane. Please evaluate how well the estimated Bayesian model can track individual hurricanes.

```{r echo=FALSE, results=FALSE}
# prior settings from MCMC.rmd, not to be shown in the pdf file.
origin_df = read.csv("hurrican703.csv")

hurricane_df = origin_df %>% 
  mutate(
    Active = ifelse(Month %in% month.name[8:10], "Active", "Inactive"),
    Active = factor(Active, levels = c("Inactive", "Active")),
    Month = factor(Month, levels = month.name[-c(2:3)]), # April-January (January ref, may choose another)
    Nature = as.factor(Nature), # TS,ET,DS,SS,NR (DS ref, may choose another)
    # note: one hurricane can have multiple natures throughout its life
    time = gsub("[()]", "", time),
    time = paste0(ifelse(substr(time, 1, 2) > 23, "19", "20"), time),
    time = as.POSIXct(time, format = "%Y-%m-%d %H:%M:%S"),
    hour = substr(time, 12, 19)
  ) %>% 
  # remove data not at six-hour time intervals. (613 observations)
  filter(hour %in% c("00:00:00", "06:00:00", "12:00:00", "18:00:00")) %>% 
  dplyr::select(-hour)

# remove hurricanes that has only 2 (<3) observations (change the threshold if you wish)
few_id = hurricane_df %>% 
  group_by(ID) %>% 
  summarize(obs = n()) %>% 
  filter(obs < 3) %>% 
  .$ID
hurricane_df = hurricane_df %>% filter(!(ID %in% few_id)) # remove 3 hurricanes

# manually correct hurricanes that have same names but are actually different
hurricane_df = 
  hurricane_df %>% 
  mutate(
    # 2 hurricanes with the name ALICE.1954
    ID = ifelse(ID == "ALICE.1954" & Month == "June", "ALICE.1954(1)", ID),
    ID = ifelse(ID == "ALICE.1954", "ALICE.1954(2)", ID),
    # 4 hurricanes with the name SUBTROP:UNNAMED.1974
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974" & Month == "June", "SUBTROP:UNNAMED.1974(1)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974" & Month == "July", "SUBTROP:UNNAMED.1974(2)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974" & Month == "August", "SUBTROP:UNNAMED.1974(3)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974", "SUBTROP:UNNAMED.1974(4)", ID),
    # 2 hurricanes with the name SUBTROP:UNNAMED.1976
    ID = ifelse(ID == "SUBTROP:UNNAMED.1976" & Month == "May", "SUBTROP:UNNAMED.1976(1)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1976", "SUBTROP:UNNAMED.1976(2)", ID)
  ) %>% 
  mutate(Season = Season - 1950) # scale the year

# data partition, no need to partition data
Training = hurricane_df

n = length(unique(Training$ID))

Y = split(Training$Wind.kt, Training$ID) %>% 
  lapply(function(x) x[-c(1:2)])

X = Training %>% 
  group_by(ID) %>% 
  slice(1) %>% 
  dplyr::select(ID, Season, Active, Nature) %>% 
  ungroup(ID)
X = model.matrix(~., X[-1])[,-1]

Z = Training %>% 
  group_by(ID) %>% 
  mutate(
    intercept = 1,
    wind_pre = lag(Wind.kt),
    lat_diff = lag(Latitude) - lag(Latitude, 2),
    long_diff = lag(Longitude) - lag(Longitude, 2),
    wind_diff = lag(Wind.kt) - lag(Wind.kt, 2),
  ) %>% 
  drop_na %>% 
  dplyr::select(ID, intercept, wind_pre, lat_diff, long_diff, wind_diff)
Z = split(Z[, names(Z)[-1]], Z$ID) %>% 
  lapply(as.matrix)

m = Training %>% 
  group_by(ID) %>% 
  summarize(obs = n() - 2) %>% 
  .$obs
```

## Prediction Functions
Import parameters from task 1 & 2.
```{r}
# load parameters
beta_list = read.csv("./data/B_list_lastmean.csv")
gamma_list = read.csv("./data/gamma_list.csv")
```

Implement the prediction process of wind speed for each hurricane in R.
```{r}
Speed_Prediction = function(beta, gamma, burn_bindex, burn_times){
  # final parameters to be used
  # the rows of beta_sample means the last 5000, 4000, 3000, 2000 and 1000
  # burn in the MC chains. change this based on the resulting plots
  # if burn-in is set to 5000, then we set burn_bindex as 1 to pick the 4th row of beta_sample
  # index is the useful samples (used for estimates & CIs)
  para_beta = beta[burn_bindex,]
  para_beta = as.matrix(para_beta)
  index = (burn_times + 1):10000
  gamma_sample = gamma[index, ]
  para_gamma = rbind(colMeans(gamma_sample))
  
  # prediction function
  Windspeed_Predict = function(index_hurricane, index_time){
    predict_speed = 
      Z[[index_hurricane]][index_time,] %*% para_beta[((index_hurricane - 1) * 5 + 1):((index_hurricane - 1) * 5 + 5)] + 
      (X %*% t(para_gamma))[index_hurricane, ]
    return(predict_speed)
  }
  
  # initialize prediction table
  Y_table = split(Training$Wind.kt, Training$ID) %>% 
    lapply(function(x) x[-c(1:2)]) %>%
    lapply(as.data.frame) %>%
    lapply(function(df) {
      df$wind_obs = df$`X[[i]]`
      df$wind_predict = df$wind_obs
      df = as.matrix(df)
      subset(df, select = c("wind_obs", "wind_predict"))
    })
  
  # updating prediction table
  for (i in 1:length(names(Z))) {
   for (j in 1:nrow(Z[[i]])) {
   Y_table[[i]][, 2][j] = Windspeed_Predict(i, j)
    } 
  }
  return(Y_table)
}
```


```{r}
Y_table_set = Speed_Prediction(beta = beta_list, gamma = gamma_list, burn_bindex = 1, burn_times = 5000)

Visual_Table = function(Y_table_input){
  Y_table = Y_table_input
  hurri_res = data.frame(ID = "example",
                         RMSE = 0,
                         R_squared = 0)
  for (i in 1:length(names(Z))) {
    # calculate
    RMSE = sqrt(mean((Y_table[[i]][,1] - Y_table[[i]][,2])^2))
    # calculate R^2
    y = Y_table[[i]][,1]
    y_hat = Y_table[[i]][,2]
    mean_y = mean(y)
    SSR = sum((y_hat - y)^2)
    SST = sum((y - mean_y)^2)
    R_squared = 1 - SSR/SST
    new_row = c(names(Z)[i], RMSE, R_squared)
    hurri_res = rbind(hurri_res, new_row)
  }
  
  hurri_res = hurri_res[-1, ]
  hurri_res$RMSE = as.numeric(hurri_res$RMSE)
  hurricane_info = Training %>% 
    group_by(ID) %>% 
    slice(1) %>% 
    dplyr::select(ID, Season, Month, Nature) %>% 
    ungroup(ID) %>%
    mutate(
      Active = ifelse(Month %in% month.name[8:10], "Active", "Inactive"),
      Active = factor(Active, levels = c("Inactive", "Active")))
  
  hurricane_loc = Training %>%
    distinct(ID, .keep_all = TRUE) %>%
    dplyr::select(ID, Latitude, Longitude, Wind.kt) %>%
    mutate(Start_Lat = Latitude,
           Start_Lon = Longitude,
           Start_Speed = Wind.kt) %>%
    dplyr::select(ID, Start_Lat, Start_Lon, Start_Speed)
  
  hurricane_info = left_join(hurricane_info, hurricane_loc, by = "ID")
  
  hurri_res = left_join(hurri_res, hurricane_info, by = "ID")
  
  hurri_res$R_squared = as.numeric(hurri_res$R_squared)
  
  return(hurri_res)
}
```

## Visualization Analysis

### Summary Table
```{r}
hurri_res = Visual_Table(Y_table_set)

hurri_res_brief = hurri_res %>%
  dplyr::select(c(1:3)) %>%
  head(15)

write.csv(hurri_res_brief, "data/prediction_brief.csv", row.names = FALSE)
write.csv(hurri_res, "data/prediction_all.csv", row.names = FALSE)

hurri_res_brief %>%
  knitr::kable(digits = 4)
```


### Overall RMSE prediction performance.
```{r}
# overall density of RMSE
ggplot(hurri_res, aes(x = RMSE)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.25, alpha = 0.8, color = "grey") +
  geom_density(size = 0.5, color = "blue", lty = 1) + 
  scale_x_continuous(breaks = seq(0, 20, by = 2)) + 
  labs(title = "Histogram of Wind Speed RMSE of Different Hurricanes", x = "Hurricane Wind Speed RMSE", y = "Density") +
  theme(plot.title = element_text(hjust = 0.5))
```


### Overall R-squared performance.
```{r warning = FALSE}
# overall density of R-squared
ggplot(hurri_res %>% filter(0 <= R_squared & R_squared <= 1), aes(x = R_squared)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.01, alpha = 0.8, color = "grey") +
  geom_density(size = 0.5, color = "blue", lty = 1) + 
  labs(title = "Histogram of Wind Speed R-squared of Different Hurricanes", x = "Hurricane Wind Speed R-squared", y = "Density") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Difference of overall RMSE using different numbers of burn-in.

Result shows that there is not obvious difference here.
```{r include = FALSE}
# density of RMSE using different burn-in times
Y_table_5000 = Speed_Prediction(beta = beta_list, gamma = gamma_list, burn_bindex = 1, burn_times = 5000)
Y_table_6000 = Speed_Prediction(beta = beta_list, gamma = gamma_list, burn_bindex = 2, burn_times = 6000)
Y_table_7000 = Speed_Prediction(beta = beta_list, gamma = gamma_list, burn_bindex = 3, burn_times = 7000)
Y_table_8000 = Speed_Prediction(beta = beta_list, gamma = gamma_list, burn_bindex = 4, burn_times = 8000)
Y_table_9000 = Speed_Prediction(beta = beta_list, gamma = gamma_list, burn_bindex = 5, burn_times = 9000)

RMSE_5000 = Visual_Table(Y_table_5000) %>%
  mutate(group = "5000")
RMSE_6000 = Visual_Table(Y_table_6000) %>%
  mutate(group = "6000")
RMSE_7000 = Visual_Table(Y_table_7000) %>%
  mutate(group = "7000")
RMSE_8000 = Visual_Table(Y_table_8000) %>%
  mutate(group = "8000")
RMSE_9000 = Visual_Table(Y_table_9000) %>%
  mutate(group = "9000")

ggplot(RMSE_5000, aes(x = RMSE)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.25, alpha = 0.8, color = "grey") +
  geom_density(size = 0.5, color = "blue", lty = 1) + 
  labs(title = "Histogram of Wind Speed RMSE using Burn-in = 5000", x = "Hurricane Wind Speed RMSE", y = "Density") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(RMSE_6000, aes(x = RMSE)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.25, alpha = 0.8, color = "grey") +
  geom_density(size = 0.5, color = "blue", lty = 1) + 
  scale_x_continuous(breaks = seq(0, 20, by = 2)) + 
  labs(title = "Histogram of Wind Speed RMSE using Burn-in = 6000", x = "Hurricane Wind Speed RMSE", y = "Density") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(RMSE_7000, aes(x = RMSE)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.25, alpha = 0.8, color = "grey") +
  geom_density(size = 0.5, color = "blue", lty = 1) + 
  scale_x_continuous(breaks = seq(0, 20, by = 2)) + 
  labs(title = "Histogram of Wind Speed RMSE using Burn-in = 7000", x = "Hurricane Wind Speed RMSE", y = "Density") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(RMSE_8000, aes(x = RMSE)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.25, alpha = 0.8, color = "grey") +
  geom_density(size = 0.5, color = "blue", lty = 1) + 
  scale_x_continuous(breaks = seq(0, 20, by = 2)) + 
  labs(title = "Histogram of Wind Speed RMSE using Burn-in = 8000", x = "Hurricane Wind Speed RMSE", y = "Density") +
  theme(plot.title = element_text(hjust = 0.5))

ggplot(RMSE_9000, aes(x = RMSE)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.25, alpha = 0.8, color = "grey") +
  geom_density(size = 0.5, color = "blue", lty = 1) + 
  scale_x_continuous(breaks = seq(0, 20, by = 2)) + 
  labs(title = "Histogram of Wind Speed RMSE using Burn-in = 9000", x = "Hurricane Wind Speed RMSE", y = "Density") +
  theme(plot.title = element_text(hjust = 0.5))
```

### The difference of RMSE distribution of difference properties.
```{r}
# distribution of RMSE on hurricane nature
nature1=ggplot(hurri_res, aes(x = Nature, y = RMSE, fill = Nature)) + 
  geom_boxplot() + 
  scale_color_manual(values = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")) +
  labs(title = "RMSE vs. Nature", x = "Nature", y = "RMSE") + 
  theme(plot.title = element_text(hjust = 0.5))


nature2=ggplot(hurri_res, aes(x = RMSE, fill = Nature)) + 
  geom_density(alpha = 0.5) + 
  labs(title = "RMSE vs. Nature", x = "RMSE", fill = "Nature") + 
  theme(plot.title = element_text(hjust = 0.5))




# distribution of RMSE on hurricane active months
ggplot(hurri_res, aes(x = Active, y = RMSE, fill = Active)) + 
  geom_boxplot() + 
  labs(title = "RMSE vs. Active/Inactive Month", x = "Active month/Inactive month", y = "RMSE") + 
  theme(plot.title = element_text(hjust = 0.5))

ggplot(hurri_res, aes(x = RMSE, fill = Active)) + 
  geom_density(alpha = 0.5) + 
  labs(title = "RMSE vs. Active/Inactive Month", x = "RMSE", fill = "Active month/Inactive month") + 
  theme(plot.title = element_text(hjust = 0.5))

ggplot(hurri_res, aes(x = Month, y = RMSE, fill = Month)) + 
  geom_boxplot() + 
  labs(title = "RMSE vs. Month of Year", x = "Month of Year", y = "RMSE") + 
  theme(plot.title = element_text(hjust = 0.5))

ggplot(hurri_res, aes(x = RMSE, fill = Month)) + 
  geom_density(alpha = 0.5) + 
  labs(title = "RMSE vs. Month of Year",x = "RMSE", fill = "Month of Year") + 
  theme(plot.title = element_text(hjust = 0.5))


# distribution of RMSE on hurricane season
ggplot(hurri_res %>% filter(Season < 20) %>% mutate(Season = factor(Season)), aes(x = RMSE, y = Season)) + 
  geom_density_ridges(scale = 0.8) + 
  labs(title = "RMSE from start year to 1970",x = "RMSE", y = "Season") + 
  theme(plot.title = element_text(hjust = 0.5))

ggplot(hurri_res %>% filter(Season >= 20 & Season < 40) %>% mutate(Season = factor(Season)), aes(x = RMSE, y = Season)) + 
  geom_density_ridges(scale = 0.8) + 
  labs(title = "RMSE from 1970 to 1990", x = "RMSE", y = "Season") + 
  theme(plot.title = element_text(hjust = 0.5))

ggplot(hurri_res %>% filter(Season >= 40 & Season < 63) %>% mutate(Season = factor(Season)), aes(x = RMSE, y = Season)) + 
  geom_density_ridges(scale = 0.8) + 
  labs(title = "RMSE from 1990 to 2013", x = "RMSE", y = "Season") + 
  theme(plot.title = element_text(hjust = 0.5))
```

### Relation between RMSE and the start location information.
(not finally included since there is not much useful information in there)
```{r eval = FALSE, include = FALSE}
# distribution of RMSE between start information
ggplot(hurri_res, aes(x = Start_Lon, y = Start_Lat, color = RMSE)) +
  geom_point(size = 3) +
  scale_color_gradient(low = "#ADD8E6", high = "#000080") +
  labs(x = "Longitude", y = "Latitude", color = "RMSE")

ggplot(hurri_res, aes(x = Start_Speed, y = RMSE, color = RMSE)) +
  geom_point(size = 3) +
  scale_color_gradient(low = "#ADD8E6", high = "#000080") +
  labs(x = "Start_Speed", y = "RMSE", color = "RMSE")

ggplot(hurri_res %>% mutate(Start_Speed = factor(Start_Speed)), aes(x = Start_Speed, y = RMSE, fill = Start_Speed)) + 
  geom_boxplot() + 
  labs(x = "Start Speed", y = "RMSE")

ggplot(hurri_res %>% mutate(Start_Speed = factor(Start_Speed)), aes(x = RMSE, fill = Start_Speed)) + 
  geom_density(alpha = 0.5) + 
  labs(x = "RMSE", fill = "Start Speed")
```


### Prediction performance on some specific example hurricanes.
```{r}
for (i in 1:5) {
  example_hurri = as.data.frame(Y_table_set[[i*5]])
  example_hurri$index = 1:nrow(example_hurri)
  
  # example visualization of example hurricanes
  graph_a = ggplot(example_hurri, aes(x = wind_predict, y = wind_obs)) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE) + 
    labs(title = paste("Observation vs. Prediction of", "Example Hurricane", i), x = "Prediction", y = "Observation") +
    theme(plot.title = element_text(hjust = 0.5))
  
  graph_b = ggplot(example_hurri, aes(x = index)) +
    geom_point(aes(y = wind_obs, color = "Observed")) +
    geom_point(aes(y = wind_predict, color = "Predicted")) +
    geom_line(aes(y = wind_obs, color = "Observed")) +
    geom_line(aes(y = wind_predict, color = "Predicted")) +
    labs(title = paste("Fitting Performance", "Example Hurricane", i), x = "Time Index", y = "Wind Speed") +
    scale_color_manual(name = "Legend",
                       values = c("Observed" = "#1E90FF", "Predicted" = "orange"),
                       labels = c("Observation", "Prediction")) + 
    theme(plot.title = element_text(hjust = 0.5))
  
  ggsave(filename = paste0("predict_plots/", "plot_", i, "_a.png"), plot = graph_a, width = 6, height = 4, dpi = 300)
  ggsave(filename = paste0("predict_plots/", "plot_", i, "_b.png"), plot = graph_b, width = 6, height = 4, dpi = 300)

}


```

```{r}
ID_name=as.matrix(Training%>% 
  group_by(ID) %>% 
  slice(1) %>% dplyr::select(ID))


plot_list_a = list()
plot_list_b = list()

for (i in 1:5) {
  example_hurri = as.data.frame(Y_table_set[[i*5]])
  example_hurri$index = 1:nrow(example_hurri)
  
  # example visualization of example hurricanes
  graph_a = ggplot(example_hurri, aes(x = wind_predict, y = wind_obs)) + 
    geom_point() + 
    geom_smooth(method = "lm", se = FALSE) + 
    labs(title = ID_name[i*5,], x = "Prediction", y = "Observation") +
    theme(plot.title = element_text(hjust = 0.5))
  
  graph_b = ggplot(example_hurri, aes(x = index)) +
    geom_point(aes(y = wind_obs, color = "Observed")) +
    geom_point(aes(y = wind_predict, color = "Predicted")) +
    geom_line(aes(y = wind_obs, color = "Observed")) +
    geom_line(aes(y = wind_predict, color = "Predicted")) +
    labs(title = ID_name[i*5,], x = "Time Index", y = "Wind Speed") +
        scale_color_manual(name = "Legend",
                       values = c("Observed" = "#1E90FF", "Predicted" = "orange"),
                       labels = c("Observation", "Prediction"))+
    theme(plot.title = element_text(hjust = 0.5))
  plot_list_a[[i]] = graph_a
  plot_list_b[[i]] = graph_b

}



grid_a = grid.arrange(grobs = plot_list_a[c(1:4)], ncol = 2, nrow=2)
grid_b = grid.arrange(grobs = plot_list_b[c(1:4)], ncol = 2, nrow=2)




ggsave(filename = "predic_plots/a_1_4.png",grid_a,width = 6, height = 4, dpi = 300)
ggsave(filename = "predic_plots/b_1_4.png",grid_b,width = 6, height = 4, dpi = 300)
```

```{r}
library(cowplot)

# create the plots
plot_list_c = list()

for (i in 1:5) {
  example_hurri = as.data.frame(Y_table_set[[i*5]])
  example_hurri$index = 1:nrow(example_hurri)
  
  graph_b = ggplot(example_hurri, aes(x = index)) +
    geom_point(aes(y = wind_obs, color = "Observed")) +
    geom_point(aes(y = wind_predict, color = "Predicted")) +
    geom_line(aes(y = wind_obs, color = "Observed")) +
    geom_line(aes(y = wind_predict, color = "Predicted")) +
    labs(title = ID_name[i*5,], x = "Time Index", y = "Wind Speed") +
    theme(legend.position = "none") +
    theme(plot.title = element_text(hjust = 0.5))
  
  plot_list_c[[i]] = graph_b
}

# arrange the plots using grid.arrange
grid_c = grid.arrange(grobs = plot_list_c[c(1:4)], ncol = 2, nrow=2)

# extract the legends from each plot


# merge the legends into one common legend
legend = cowplot::get_legend(plot_list_c[[1]]+theme(legend.position="bottom"))

# add the common legend to the grid of plots
plot_grid(grid_c, legend, ncol = 1, rel_heights = c(1, .1))

```

