---
title: "Task 4"
output: 
    pdf_document
---

```{r include=FALSE, echo=FALSE}
library(MASS)
library(tidyverse)
```

**Task 2:** With the estimated model parameters and covariate values, you can calculate the predicted wind speed for each time point using the model equation. This way, you can track the hurricane and compare the predicted wind speeds with the actual wind speeds recorded during the hurricane. Please evaluate how well the estimated Bayesian model can track individual hurricanes.

```{r echo=FALSE, results=FALSE}
# prior settings from MCMC.rmd, not to be shown in the pdf file.
origin_df = read.csv("hurrican703.csv")

hurricane_df = origin_df %>% 
  mutate(
    Month = factor(Month, levels = month.name[-c(2:3)]), # April-January (January ref, may choose another)
    Nature = as.factor(Nature), # TS,ET,DS,SS,NR (DS ref, may choose another)
    # note: one hurricane can have multiple natures throughout its life
    time = gsub("[()]", "", time),
    time = paste0(ifelse(substr(time, 1, 2) > 23, "19", "20"), time),
    time = as.POSIXct(time, format = "%Y-%m-%d %H:%M:%S"),
    hour = substr(time, 12, 19)
  ) %>% 
  # remove data not at six-hour time intervals. (613 observations)
  filter(hour %in% c("00:00:00", "06:00:00", "12:00:00", "18:00:00")) %>% 
  dplyr::select(-hour)

# remove hurricanes that has only 2 (<3) observations (change the threshold if you wish)
few_id = hurricane_df %>% 
  group_by(ID) %>% 
  summarize(obs = n()) %>% 
  filter(obs < 3) %>% 
  .$ID
hurricane_df = hurricane_df %>% filter(!(ID %in% few_id)) # remove 3 hurricanes

# manually correct hurricanes that have same names but are actually different
hurricane_df = 
  hurricane_df %>% 
  mutate(
    # 2 hurricanes with the name ALICE.1954
    ID = ifelse(ID == "ALICE.1954" & Month == "June", "ALICE.1954(1)", ID),
    ID = ifelse(ID == "ALICE.1954", "ALICE.1954(2)", ID),
    # 4 hurricanes with the name SUBTROP:UNNAMED.1974
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974" & Month == "June", "SUBTROP:UNNAMED.1974(1)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974" & Month == "July", "SUBTROP:UNNAMED.1974(2)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974" & Month == "August", "SUBTROP:UNNAMED.1974(3)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1974", "SUBTROP:UNNAMED.1974(4)", ID),
    # 2 hurricanes with the name SUBTROP:UNNAMED.1976
    ID = ifelse(ID == "SUBTROP:UNNAMED.1976" & Month == "May", "SUBTROP:UNNAMED.1976(1)", ID),
    ID = ifelse(ID == "SUBTROP:UNNAMED.1976", "SUBTROP:UNNAMED.1976(2)", ID)
  )

# data partition, no need to partition data
Training = hurricane_df

n = length(unique(Training$ID))

Y = split(Training$Wind.kt, Training$ID) %>% 
  lapply(function(x) x[-c(1:2)])

X = Training %>% 
  group_by(ID) %>% 
  slice(1) %>% 
  dplyr::select(ID, Season, Month, Nature) %>% 
  ungroup(ID)
X = model.matrix(~., X[-1])[,-1]

Z = Training %>% 
  group_by(ID) %>% 
  mutate(
    intercept = 1,
    wind_pre = lag(Wind.kt),
    lat_diff = lag(Latitude) - lag(Latitude, 2),
    long_diff = lag(Longitude) - lag(Longitude, 2),
    wind_diff = lag(Wind.kt) - lag(Wind.kt, 2),
  ) %>% 
  drop_na %>% 
  dplyr::select(ID, intercept, wind_pre, lat_diff, long_diff, wind_diff)
Z = split(Z[, names(Z)[-1]], Z$ID) %>% 
  lapply(as.matrix)

m = Training %>% 
  group_by(ID) %>% 
  summarize(obs = n() - 2) %>% 
  .$obs
```

### Solution

Import parameters from task 1 & 2.
```{r}
# load parameters
beta_list = read.csv("./data/B_list_partial.csv")
mu_list = read.csv("./data/mu_list.csv")
Sigma_list = read.csv("./data/Sigma_matrix_list.csv")
gamma_list = read.csv("./data/gamma_list.csv")
sigma_list = read.csv("./data/sigma_list.csv")$x

# burn in the MC chains. change this based on the resulting plots
burn = 500 
# index of useful samples (used for estimates & CIs)
index = (burn + 1):10000
beta_sample = beta_list[index,]
mu_sample = mu_list[index,]
gamma_sample = gamma_list[index,]
```

Implement the prediction process of wind speed for each hurricane in R.
```{r}
# final parameters to be used
para_beta = rbind(colMeans(beta_sample))
para_mu = rbind(colMeans(mu_sample))
para_gamma = rbind(colMeans(gamma_sample))

# prediction function
Windspeed_Predict = function(index_hurricane, index_time){
  predict_speed = 
    Z[[index_hurricane]][index_time,] %*% para_beta[index_hurricane:(index_hurricane + 4)] + 
    (X %*% t(para_gamma))[index_hurricane, ]
  return(predict_speed)
}

# functional test
Windspeed_Predict(1, 26)

# initialize prediction table
Y_table = split(Training$Wind.kt, Training$ID) %>% 
  lapply(function(x) x[-c(1:2)]) %>%
  lapply(as.data.frame) %>%
  lapply(function(df) {
    df$wind_true = df$`X[[i]]`
    df$wind_predict = df$wind_true
    df = as.matrix(df)
    subset(df, select = c("wind_true", "wind_predict"))
  })

# updating prediction table
for (i in 1:length(names(Z))) {
 for (j in 1:nrow(Z[[i]])) {
 Y_table[[i]][, 2][j] = Windspeed_Predict(i, j)
  } 
}
```

Data visualization for prediction results.
```{r}

```

